|

### 2019-03-17 Paper.Rmd
more paper edits

---
title: "ECON 21320: Final Paper"
author: "Eric Milstein and Tak Hogan"
date: "March 21, 2019"
bibliography: references.bib
output: pdf_document
---

@@

# Abstract
[@aliprantisBlowingItKnocking2015]

@@

Few empirical studies have been conducted into the effects of inclusionary housing policies. While there have been several policy reports and feasibility/cost-benefit analyses, there is little academic literature. Some of the potential barriers to research include the small scope of these programs, the recency of their implementation, a lack of available data, and legal complexities in evaluating the effects of these policies. However, there is related literature on other “place-based” policies (enterprise zones, neighborhood revitalization projects, etc.), spatial econometrics and spillover effects, rent control, and other affordable housing programs.

@@

# Methods

@@

@@

Due to the difficulties in estimating this model, we have considered several related research questions. Another potential study that we discussed is to evaluate whether developers are acting rationally in choosing to pay the in-lieu fee. If we were to pursue this question, we might construct some estimate of a developer’s “expected payoff” by observing trends in housing prices/values in surrounding areas. We might then regress the expected payoff on a “paid in-lieu fee” dummy to determine if they made the right decision in choosing whether or not to pay the fee.
Data

# Data and Analysis

@@

```{r}
# 
# # Summary statistics table
# 
# our_summary1 <-
#   list("Assessed Land Value" =
#          list("min" = ~ min(total.clean$Assessed.Land.Value),
#               "median" = ~ median(total.clean$Assessed.Land.Value),
#               "max" = ~ max(total.clean$Assessed.Land.Value),
#               "mean (sd)" = ~ qwraps2::mean_sd(total.clean$Assessed.Land.Value)),
#        "Assessed Improvement Value" =
#          list("min" = ~ min(total.clean$Assessed.Improvement.Value),
#               "median" = ~ median(total.clean$Assessed.Improvement.Value),
#               "max" = ~ max(total.clean$Assessed.Improvement.Value),
#               "mean (sd)" = ~ qwraps2::mean_sd(total.clean$Assessed.Improvement.Value))
#               "Forward Gears" =
#        list("Three" = ~ qwraps2::n_perc0(total.clean$gear == 3),
#             "Four"  = ~ qwraps2::n_perc0(total.clean$gear == 4),
#             "Five"  = ~ qwraps2::n_perc0(total.clean$gear == 5))
#        )
# 
# ### Overall
# by_415 <- summary_table(total.clean, our_summary1)
# kable(by_415)

```

>>>>

---
title: |
  | \LARGE{The Effect of Affordable Housing on Neighboring Property Values: }
  | \LARGE{Evidence from San Francisco's Inclusionary Housing Program \vspace{0.5cm}}
author: |
  | Eric Milstein and Tak Hogan
  | Applications of Econometric and Data Science Methods
date: "March 21, 2019"
bibliography: references.bib
output: pdf_document
fontsize: 11pt
fig_caption: yes
indent: true
geometry: margin=1in
header-includes:
- \usepackage{placeins}
- \usepackage{setspace}
- \usepackage{chngcntr}
- \onehalfspacing

abstract: "This paper evaluates San Francisco's Inclusionary Housing Program by estimating the impact of affordable developments on assessed and predicted property values in the surrounding area. We use a unique panel dataset constructed from parcel-level tax assessments, American Community Survey microdata, and price estimates scraped from real estate listing service Zillow. Exploiting the city's mandate that all new developments either designate affordable units or pay an in-lieu fee, we use a difference-in-differences design to compare changes in surrounding property values between these two groups. We find that on average, property values _____ by _____ in areas that compared to those that ."

---

@@

@@

### San Francisco's Inclusionary Housing Program

@@

Few empirical studies have been conducted into the effects of inclusionary housing policies. While there have been several policy reports and feasibility/cost-benefit analyses, there is little academic literature. Some of the potential barriers to research include the small scope of these programs, the recency of their implementation, a lack of available data, and legal complexities in evaluating the effects of these policies. However, there is related literature on other “place-based” policies (enterprise zones, neighborhood revitalization projects, etc.), spatial econometrics and spillover effects, rent control, and other affordable housing programs.
[@aliprantisBlowingItKnocking2015]

# Research Design

@@

### Model

$$\text{Y}_{it}=\alpha+\beta S_{i}+\gamma T_{t}+\delta(S_{i}\times T_{t})+\rho\mathbf{X}_{it}+\epsilon_{it}$$

$$
\begin{cases}
Y_{it} & \text{Median Property Values in Surrounding Area }\\
 & \text{at Time }t\\
S_{i} & \text{Section 415 Declaration Dummy (Affordable vs. Pay Fee)}\\
T_{t} & \text{Pre vs. Post Construction Dummy}\\
\mathbf{X}_{it} & \text{Vector of Covariates}
\end{cases}
$$

The difference-in-differences estimate:

$$\hat{\delta}_{DD}=\bar{Y}_{\text{post}}^{\text{affordable}}-\bar{Y}_{\text{pre}}^{\text{affordable}}-(\bar{Y}_{\text{post}}^{\text{pay fee}}-\bar{Y}_{\text{pre}}^{\text{pay fee}})$$

Interpreting coefficients

$$\text{Y}_{it}=\alpha+\beta S_{i}+\gamma T_{t}+\delta(S_{i}\times T_{t})+\rho\mathbf{X}_{it}+\epsilon_{it}$$

$$\begin{cases}
\alpha & \text{Average pre-construction median surrounding property value }\\
 & \text{for non-affordable properties}\\
\beta & \text{Affordable housing-specific effect on surrounding property values }\\
\gamma & \text{Time trend common to both affordable and non-affordable properties }\\
\delta & \text{Diff-in-diff estimator: true effect of treatment}
\end{cases}$$

Assumption underlying diff-in-diff estimation: in the absence of affordable housing, property i's outcome at time t is given by:

$$E[Y_{i}|S_{i}=0,t=\tau]=\gamma_{i}+\lambda_{\tau}$$

Two implicit identifying assumptions:

• Selection bias relates to fixed characteristics of properties ($\gamma_{i}$)

• Time trend ($\lambda_{\tau}$) same for affordable and non-affordable properties

@@

# Data Description

@@

###

# Analysis
